--- 
title: "Movement Mapper"
format: dashboard
server: shiny
execute:
  daemon: false
  freeze: false
---

```{r}
#| context: setup
library(sf)
library(amt)
library(bslib)
library(leaflet)
library(tidyverse)
library(DT)
options(shiny.maxRequestSize=100*1024^2) 
```

## {.sidebar}

**Upload data**

```{r}
fileInput("csv1", "Movement data (csv):", accept=".csv")
fileInput("csv2", "Segments table (csv):", accept=".csv")
hr()
selectInput("id", "Select caribou:", choices=NULL)
selectInput("year", "Select year:", choices=NULL)
checkboxInput("extend", "Extend fall season into next year", value = FALSE)
textInput("spring_notes", "Spring notes:", value=NULL)
textInput("fall_notes", "Fall notes:", value=NULL)
hr()
```

**NSD vs Time plots**

The two plots on the right show the relationship between NSD and "Day of year". NSD calculates the net squared displacement (i.e., the squared distance from the first location of a track) for a track.

```{r}
hr()
actionButton("log", "Save segments")
br()
hr()
downloadButton("save", "Download final segments")
```

# Plots {orientation="columns" scrolling="true"}

## Column {width=30%}

### Row {height="50%"}

```{r}
tableOutput("tab1a")
leafletOutput("map1")
```

### Row {height="50%"}

```{r}
tableOutput("tab1b")
leafletOutput("map2")
```

```{r}
#| context: server
observeEvent(input$csv1, {
  ids <- as.character(sort(unique(gps()$id)))
  yrs <- as.character(sort(unique(gps()$year)))
  updateSelectInput(session, "id", choices=ids, selected=ids[1])
  updateSelectInput(session, "year", choices=yrs, selected=yrs[1])
})


gps <- eventReactive(input$csv1, {
  req(input$csv1)
  read_csv(input$csv1$datapath) |>
    mutate(year = year(timestamp), yday=yday(timestamp)) # |> filter(id %in% 43140:43146)
})

segments_rv <- reactiveVal(NULL)

mig <- eventReactive(input$csv2, {
  if (is.null(input$csv2)) {
    tibble::tibble(
      id = integer(),
      year = integer(),
      migration = character(),
      start = character(),
      end = character(),
      notes = character()
    )
  } else {
    readr::read_csv(input$csv2$datapath, col_types = cols(
        id = col_integer(),
        year = col_integer(),
        migration = col_character(),
        start = col_character(),
        end = col_character(),
        notes = col_character()
      ))
  }
}, ignoreNULL = FALSE)

# Initialize segments when upload happens
observeEvent(mig(), {
  segments_rv(mig())
})

observe({
  note_spring <- mig()$notes[mig()$id==input$id & mig()$year==input$year & mig()$migration=="Spring"]
  note_fall <- mig()$notes[mig()$id==input$id & mig()$year==input$year & mig()$migration=="Fall"]
  updateTextInput(session, "spring_notes", value=note_spring)
  updateTextInput(session, "fall_notes", value=note_fall)
})

observeEvent(input$csv2, {
  start_spring <- yday(mig()$start[mig()$id==input$id & mig()$year==input$year & mig()$migration=="Spring"])
  end_spring <- yday(mig()$end[mig()$id==input$id & mig()$year==input$year & mig()$migration=="Spring"])
  start_fall <- yday(mig()$start[mig()$id==input$id & mig()$year==input$year & mig()$migration=="Fall"])
  end_fall <- yday(mig()$end[mig()$id==input$id & mig()$year==input$year & mig()$migration=="Fall"])
  updateSliderInput(session, "spring", min=1, max=365, value=c(start_spring,end_spring), step=1)
  updateSliderInput(session, "fall", min=1, max=365, value=c(start_fall,end_fall), step=1)
})

observe({
  start_spring <- yday(mig()$start[mig()$id==input$id & mig()$year==input$year & mig()$migration=="Spring"])
  end_spring <- yday(mig()$end[mig()$id==input$id & mig()$year==input$year & mig()$migration=="Spring"])
  start_fall <- yday(mig()$start[mig()$id==input$id & mig()$year==input$year & mig()$migration=="Fall"])
  end_fall <- yday(mig()$end[mig()$id==input$id & mig()$year==input$year & mig()$migration=="Fall"])
  updateSliderInput(session, "spring", min=1, max=365, value=c(start_spring,end_spring), step=1)
  updateSliderInput(session, "fall", min=1, max=365, value=c(start_fall,end_fall), step=1)
})

observeEvent(input$extend, {
  start_spring <- yday(mig()$start[mig()$id==input$id & mig()$year==input$year & mig()$migration=="Spring"])
  end_spring <- yday(mig()$end[mig()$id==input$id & mig()$year==input$year & mig()$migration=="Spring"])
  start_fall <- yday(mig()$start[mig()$id==input$id & mig()$year==input$year & mig()$migration=="Fall"])
  end_fall <- yday(mig()$end[mig()$id==input$id & mig()$year==input$year & mig()$migration=="Fall"])
  updateSliderInput(session, "spring", min=1, max=426, value=c(start_spring,end_spring), step=1)
  updateSliderInput(session, "fall", min=1, max=426, value=c(start_fall,end_fall), step=1)
})

trk <- eventReactive(input$csv1, {
  gps() |>
    make_track(.x=longitude, .y=latitude, .t=timestamp, id=id, longitude=longitude, 
      latitude=latitude, elevation=elevation, year=year, yday=yday, crs = 4326) 
})

Boundary <- eventReactive(input$csv1, {
  req(trk())
  hr_kde(trk(), levels=0.9999) |> hr_isopleths()
})

trk_select <- reactive({
  req(input$year)
  if (input$extend) {
    # keep selected year + next year
    trk <- trk() |> 
      filter(id==input$id, year %in% c(as.numeric(input$year), as.numeric(input$year) + 1))
  } else {
    # keep only selected year
    trk <- trk() |> 
      filter(id==input$id, year == input$year)
  }

  trk <- trk |> 
    arrange("t_") |> 
    mutate(plot_yday = ifelse(year != input$year, yday + 365, yday)) |> 
      filter(plot_yday < 427)
  trk <- trk |> mutate(nsd=nsd(trk))
  trk
})

last_day <- reactive(ifelse(input$extend, 426, 365))

# Initialize spring map 
output$map1 <- renderLeaflet({
  leaflet(Boundary(), options = leafletOptions(attributionControl = FALSE)) |>
    addProviderTiles("Esri.WorldImagery", group="Esri.WorldImagery") |>
    addProviderTiles("Esri.WorldGrayCanvas", group="Esri.WorldGrayCanvas") |>
    addProviderTiles("Esri.WorldTopoMap", group="Esri.WorldTopoMap") |>
    addPolygons(fillOpacity=0, stroke=TRUE, color="black", weight=2, group="Boundary") |>
    addLayersControl(
      position = "topright",
      baseGroups = c("Esri.WorldTopoMap", "Esri.WorldImagery", "Esri.WorldGrayCanvas"),
      overlayGroups = c("Boundary", "GPS", "GPS Spring", "Tracks"),
      options = layersControlOptions(collapsed = TRUE)
    )
})

# Update the spring layers
observe({
  req(trk_select())
  lines <- trk_select()
  spring <- lines |> filter(plot_yday >= input$spring[1] & plot_yday <= input$spring[2])
  leafletProxy("map1") |>
    clearGroup("GPS") |>
    clearGroup("GPS Spring") |>
    clearGroup("Tracks") |>
    addCircles(data = lines, ~x_, ~y_, color="black", fill="black", weight=1, radius=1, group="GPS") |>
    addCircles(data = spring, ~x_, ~y_, color="red", fill="red", weight=10, radius=10, group="GPS Spring") |>
    addPolylines(data = lines, ~x_, ~y_, stroke=TRUE, color="black", weight=1, group="Tracks")
})


# Initialize fall map 
output$map2 <- renderLeaflet({
  leaflet(Boundary(), options = leafletOptions(attributionControl = FALSE)) |>
    addProviderTiles("Esri.WorldImagery", group="Esri.WorldImagery") |>
    addProviderTiles("Esri.WorldGrayCanvas", group="Esri.WorldGrayCanvas") |>
    addProviderTiles("Esri.WorldTopoMap", group="Esri.WorldTopoMap") |>
    addPolygons(fillOpacity=0, stroke=TRUE, color="black", weight=2, group="Boundary") |>
    addLayersControl(
      position = "topright",
      baseGroups = c("Esri.WorldTopoMap", "Esri.WorldImagery", "Esri.WorldGrayCanvas"),
      overlayGroups = c("Boundary", "GPS", "GPS Fall", "Tracks"),
      options = layersControlOptions(collapsed = TRUE)
    )
})

# Update the fall layers
observe({
  req(trk_select())
  lines <- trk_select()
  fall   <- lines |> filter(plot_yday >= input$fall[1] & plot_yday <= input$fall[2])
  leafletProxy("map2") |>
    clearGroup("GPS") |>
    clearGroup("GPS Fall") |>
    clearGroup("Tracks") |>
    addCircles(data = lines, ~x_, ~y_, color="black", fill="black", weight=1, radius=1, group="GPS") |>
    addCircles(data = fall, ~x_, ~y_, color="red", fill="red", weight=10, radius=10, group="GPS Fall") |>
    addPolylines(data = lines, ~x_, ~y_, stroke=TRUE, color="black", weight=1, group="Tracks")
})

```

```{r}
#| context: server
oneRow <- reactive({
  # Subtracts 1 because DOY 1 corresponds to January 1st
  start_spring <- as.character(as.Date(input$spring[1]-1, origin = paste0(input$year, "-01-01", format="%b-%d")))
  end_spring <- as.character(as.Date(input$spring[2]-1, origin = paste0(input$year, "-01-01", format="%b-%d")))
  start_fall <- as.character(as.Date(input$fall[1]-1, origin = paste0(input$year, "-01-01", format="%b-%d")))
  end_fall <- as.character(as.Date(input$fall[2]-1, origin = paste0(input$year, "-01-01", format="%b-%d")))
  v <- c(start_spring, end_spring, start_fall, end_fall)
})

output$tab1a <- renderTable({ 
  req(input$csv1)
  req(input$id)
  req(input$year)
  if (!is.null(input$csv1)) {
    x <- tibble(
      id=c(input$id,input$id),
      year=c(input$year,input$year),
      migration=c("Spring","Fall"),
      start=c(oneRow()[1], oneRow()[3]), 
      end=c(oneRow()[2], oneRow()[4]))
    x[1,]
  }
})

output$tab1b <- renderTable({ 
  req(input$csv1)
  req(input$id)
  req(input$year)
  if (!is.null(input$csv1)) {
    x <- tibble(
      id=c(input$id,input$id),
      year=c(input$year,input$year),
      migration=c("Spring","Fall"),
      start=c(oneRow()[1], oneRow()[3]), 
      end=c(oneRow()[2], oneRow()[4]))
    x[2,]
  }
})

observeEvent(input$log, {
  req(segments_rv())
  browser()
  updateActionButton(session, "log", label = "Saved!")
  
  dat <- segments_rv()
  
  # Decide start/end values based on checkboxes
  start_spring <- if (isTRUE(input$use_spring)) oneRow()[1] else NA_character_
  end_spring   <- if (isTRUE(input$use_spring)) oneRow()[2] else NA_character_
  start_fall   <- if (isTRUE(input$use_fall)) oneRow()[3] else NA_character_
  end_fall     <- if (isTRUE(input$use_fall)) oneRow()[4] else NA_character_
  
  new_entries <- tibble::tibble(
    id        = as.numeric(input$id),
    year      = as.numeric(input$year),
    migration = c("Spring", "Fall"),
    start     = c(start_spring, start_fall),
    end       = c(end_spring, end_fall),
    notes      = c(
      ifelse(input$spring_notes == "", NA_character_, input$spring_notes),
      ifelse(input$fall_notes == "", NA_character_, input$fall_notes)
    )
  )

  # Update existing rows or append
  segments_rv(dplyr::rows_upsert(
    
    dat,
    new_entries,
    by = c("id", "year", "migration")
  ))
  later::later(function() {
      updateActionButton(session, "log", label = "Save segments")
    }, 2)
})
```

```{r}
#| context: server
# download the *latest* tibble
output$save <- downloadHandler(
  filename = function() {
    paste0("migration_segments_", Sys.Date(), ".csv")
  },
  content = function(file) {
    req(segments_rv())
    write_csv(segments_rv(), file)
  }
)
```

## Column {width=70%}

### Row {height="50%"}

```{r}
checkboxInput("use_spring", "Spring migration:", value = TRUE)
sliderInput("spring", " ", min=1, max=365, value=c(91,181), step=1, width="100%")
plotOutput("plotSpring")
```

### Row {height="50%"}

```{r}
checkboxInput("use_fall", "Fall migration:", value = TRUE)
sliderInput("fall", " ", min=1, max=365, value=c(244,304), step=1, width="100%")
plotOutput("plotFall")
```

```{r}
#| context: server
output$plotSpring <- renderPlot({
  ggplot(trk_select()) +
    aes(y = nsd, x = plot_yday, fill = year, text = id) +
    geom_point(size = 1) + 
    geom_path(color = "black") +
    geom_vline(xintercept = input$spring[1], color = "red", linetype = "dashed", size = 1) +
    geom_vline(xintercept = input$spring[2], color = "red", linetype = "dashed", size = 1) +
    scale_fill_viridis_c("year") +
    scale_x_continuous(limits = c(1, last_day())) +
    theme(legend.position = "none") +
    labs(x = "Day of year (extended)", y = "NSD")
})

output$plotFall <- renderPlot({
ggplot(trk_select()) +
  aes(y = nsd, x = plot_yday, fill = year, text = id) +
  geom_point(size = 1) + 
  geom_path(color="black") +
  geom_vline(xintercept = input$fall[1], color = "red", linetype = "dashed", size = 1) +
  geom_vline(xintercept = input$fall[2], color = "red", linetype = "dashed", size = 1) +
  scale_fill_viridis_c("year") +
  scale_x_continuous(limits = c(1, last_day())) +
  theme(legend.position="none") +
  labs(x = "Day of year", y = "NSD")
})
```

# Tables

## Column {width=50%}

```{r}
#| context: server
output$tab1 <- renderDataTable({
  req(mig())
  req(segments_rv())

  df1 <- mig() |> 
    mutate(id = as.integer(id), 
           year = as.integer(year),
           start = ifelse(is.na(start), "NA", start),
           end = ifelse(is.na(end), "NA", end)) |> 
    mutate(` ` = row_number()) |>
    select(` `, everything())
  
  if (nrow(df1) == 0) {
    # Return empty datatable if mig() not initialized 
    return(DT::datatable(
      df1,
      caption = "Table empty",
      rownames = FALSE,
      options = list(dom = 'tip'),
      class = "compact"
    ))
  }
  
  df2 <- segments_rv() |> 
    mutate(id = as.integer(id), 
           year = as.integer(year),
           start = ifelse(is.na(start), "NA", start),
           end = ifelse(is.na(end), "NA", end)) |> 
    mutate(` ` = row_number())  |>
    select(` `, everything())
  
  # row counts
  n <- min(nrow(df1), nrow(df2))

  # row indices where they differ
  diff_rows <- which(apply(df1 != df2, 1, any))

  if(length(diff_rows)==0){
    return(DT::datatable(
      df1,
      caption = "",
      rownames = FALSE,
      options = list(dom = 'tip'),
      class = "compact"
    ))
  }else{
     return(DT::datatable(df1,
    caption = "",
    rownames = FALSE,
    options = list(dom = 'tip', scrollX = TRUE, pageLength = n),
    class = "compact"
  ) %>%
    formatStyle(
      columns = names(df1),
      target = "row",
      backgroundColor = styleEqual(diff_rows, rep("yellow", length(diff_rows)))
    )
   )
  }
})

# Table 2
output$tab2 <- renderDataTable({
  req(mig())
  req(segments_rv())

  df1 <- mig() |> 
    mutate(id = as.integer(id), 
           year = as.integer(year),
           start = ifelse(is.na(start), "NA", start),
           end = ifelse(is.na(end), "NA", end)) |> 
    mutate(` ` = row_number())  |>
    select(` `, everything())
  
  df2 <- segments_rv() |> 
    mutate(id = as.integer(id), 
           year = as.integer(year),
           start = ifelse(is.na(start), "NA", start),
           end = ifelse(is.na(end), "NA", end)) |> 
    mutate(` ` = row_number())  |>
    select(` `, everything())
  
  if (nrow(df1) == 0) {
    return(DT::datatable(
      df2,
      caption = "",
      rownames = FALSE,
      options = list(dom = 'tip'),
      class = "compact"
    ))
  } else {
    n <- min(nrow(df1), nrow(df2))
    diff_rows <- which(apply(df1 != df2, 1, any))
    
    if(length(diff_rows)==0){
      return(DT::datatable(
      df2,
      caption = "",
      rownames = FALSE,
      options = list(dom = 'tip'),
      class = "compact"
    ))
    } else {
      return(DT::datatable(
        df2,
        caption = "",
        rownames = FALSE,
        options = list(dom = 'tip', scrollX = TRUE, pageLength = nrow(df2)),
        class = "compact"
        ) %>%
        formatStyle(
          columns = names(df2),
          target = "row",
          backgroundColor = styleEqual(diff_rows, rep("yellow", length(diff_rows)))
        )
      )
    }
  }
})

```

```{r}
#| title: Original
DTOutput("tab1")
```

```{r}
#| title: Revised
DTOutput("tab2")
```

# Help

{{< include help.md >}}
